---
title: "Box-Jenkins Methodology"
author: "Carolina Vasconcelos"
date:  "`r Sys.Date()`"
output: rmdformats::readthedown
---
# Introduction

The purpose of this notebook is to apply the Box-Jenkins Methodology as a tutorial/practical exercise. This is not an example of the project itself, you should read the evaluation criteria to ensure you address all the required points.

```{r load, echo=FALSE, message=FALSE, warning=FALSE}
library(fpp3)
library(urca)
library(kableExtra)

japan_data <- aus_arrivals %>%
  filter(Origin == "Japan")

japan_data_train <- aus_arrivals %>%
  filter(Origin == "Japan") %>% 
  filter(year(Quarter) < 2011)
```

# Stabilize variance

The first step is to visually assess the time series plot. Do we need to stabilize the variance?

```{r var, echo=FALSE, message=FALSE, warning=FALSE}
japan_data_train %>%
  mutate(`log(Arrivals)` = log(Arrivals)) %>% 
  select(-Origin) %>% 
  pivot_longer(cols = -Quarter) %>% 
  ggplot()+
  geom_line(aes(x=Quarter, y=value)) +
  facet_wrap(~ name, scales = 'free_y') +
  theme_minimal() +
  labs(y='')
```

# Stationarity

The second step is to check for the presence of a unit root, using the GTS t-sig approach the results showed that it was necessary to take the seasonal and first differences.

```{r stationarity, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}
summary(ur.df(japan_data_train$Arrivals,
              type='drift', lags=10))

japan_data_train %>%
  autoplot(difference(Arrivals, 4))

summary(ur.df(na.omit(difference(japan_data_train$Arrivals,4)),
              type='drift', lags=10))

japan_data_train %>%
  autoplot(difference(Arrivals, 4) %>%  difference())

japan_data_train <- japan_data_train %>% 
              mutate(diff4_y = difference(Arrivals, 4),
                     diff_y = difference(diff4_y))

summary(ur.df(na.omit(japan_data_train$diff_y),
              type='drift', lags=10))
```

```{r figure, echo=FALSE,  message=FALSE, warning=FALSE}
japan_data_train %>%
  autoplot(difference(Arrivals, 4) %>%  difference()) +
  labs(y='')
```


# ACF and PACF plot - Identify model

The third step is identify the SARIMA model orders through the assessment of the ACF and PACF. 

```{r identify,  message=FALSE, warning=FALSE}
japan_data_train %>% 
  gg_tsdisplay(difference(Arrivals, 4) %>%  
                 difference(), plot_type = 'partial')
```

# Estimate models

The fourth step is to estimate the model and asssess the information criteria. Compare and comment.

```{r estimation, echo=FALSE, message=FALSE, warning=FALSE}
japan_data_train %>% 
  model(
    arima111011 = ARIMA(Arrivals ~ pdq(1,1,1) + PDQ(0,1,1)),
    arima211011 = ARIMA(Arrivals ~ pdq(2,1,1) + PDQ(0,1,1)),
    arima111111 = ARIMA(Arrivals ~ pdq(1,1,1) + PDQ(1,1,1)),
    arima211111 = ARIMA(Arrivals ~ pdq(2,1,1) + PDQ(1,1,1)),
    auto = ARIMA(Arrivals)
  ) %>%
  glance() %>% 
  select(-c(ar_roots, ma_roots)) %>% 
  kable()
```

# Diagnostic checks

We assess the ACF and time series plot of the residuals:
```{r residuals, echo=FALSE, message=FALSE, warning=FALSE}

fit <- japan_data_train %>% 
  model(
    arima111011 = ARIMA(Arrivals ~ pdq(1,1,1) + PDQ(0,1,1)),
    arima211011 = ARIMA(Arrivals ~ pdq(2,1,1) + PDQ(0,1,1)),
    arima111111 = ARIMA(Arrivals ~ pdq(1,1,1) + PDQ(1,1,1)),
    arima211111 = ARIMA(Arrivals ~ pdq(2,1,1) + PDQ(1,1,1)),
    auto = ARIMA(Arrivals)
  ) 

fit %>%
  select(arima111011) %>%
  gg_tsresiduals()
```

And perform some statistical tests:
```{r residuals_test, echo=FALSE, message=FALSE, warning=FALSE}
res1 = augment(fit) %>% 
        filter(.model == 'arima111011') %>% 
        features(.innov, ljung_box, lag = 11, dof=3)

res2 = augment(fit) %>% 
        filter(.model == 'arima211011') %>% 
        features(.innov, ljung_box, lag = 11, dof=4)

res3= augment(fit) %>% 
      filter(.model == 'arima111111') %>% 
      features(.innov, ljung_box, lag = 11, dof=4)

res4 = augment(fit) %>% 
        filter(.model == 'arima211111') %>% 
        features(.innov, ljung_box, lag = 11, dof=5)

bind_rows(res1, res2, res3, res4) %>% 
  kable()
```


# Check accuracy measures

We will check the accuracy measures and select a model:

```{r forecast, echo=FALSE, message=FALSE, warning=FALSE}

fit %>%
  forecast(h = 7) %>% 
  accuracy(japan_data) %>% 
  kable()
```

Then, we plot it!

```{r forecastplot, message=FALSE, warning=FALSE}

fit %>%
  forecast(h = 7) %>% 
  accuracy(japan_data) %>% 
  kable()

fit %>%
  forecast(h = 7) %>% 
  filter(.model == 'arima111011') %>% 
  autoplot(japan_data) 

```


