---
title: "Forecasting Monthly Marriages"
author: "Carlota Marto 20241729, Francisca Teixeira 20241702, Inês Caseiro "
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(forecast)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tsibble)
library(feasts)
library(fable)
library(dplyr)
library(knitr)
```

## 1. Introduction

This project analyzes and forecasts **monthly marriages in Queensland, Australia**, using the Box-Jenkins methodology (ARIMA/SARIMA).\
The dataset is sourced from the Queensland Government Open Data portal ([Marriages by Month](https://www.data.qld.gov.au/dataset/marriages-by-month)), which provides official monthly counts of marriages registered across the state.

The objective of this analysis is to identify trends and seasonal patterns in the number of marriages, fit a suitable SARIMA model, and produce reliable forecasts for the next 12 months.\
These forecasts can provide insights for government planning, service provision, and understanding social trends related to marriage in Queensland.

```{r load-data}
# Define the correct path for the CSV file
file_path <- "/Users/franciscateixeira/Documents/Forecasting_Project/marriages_queensland/combined_file.csv"

# Load the CSV file
data <- read.csv(file_path)

# Create a Date column from Year and Month
data <- data %>%
  mutate(Date = as.Date(paste(year, Month, "1", sep="-"))) %>%
  arrange(Date)

# Convert to tsibble
data_ts <- data %>%
  select(Date, Marriages) %>%
  as_tsibble(index = Date)

# Show the first 10 rows of the dataset
kable(head(data_ts, 10), caption = "First 10 rows of Monthly Marriages Data")

```

```{r }
data_ts %>%
  ggplot(aes(x = Date, y = Marriages)) +
  geom_line(color="red") +
  labs(title="Monthly Marriages Over Time", x="Date", y="Marriages") +
  theme_minimal()
```

```{r }
data %>%
  mutate(log_Marriages = log(Marriages)) %>%  # Log transform the 'Marriages' column
  pivot_longer(cols = c(Marriages, log_Marriages), names_to = "Transformation", values_to = "Value") %>%  # Reshape data to long format
  ggplot(aes(x = Date, y = Value)) +  # Plot Date vs the transformed or original 'Marriages'
  geom_line() +
  facet_wrap(~ Transformation, scales = "free_y") +  # Facet by Transformation (original vs log-transformed)
  theme_minimal() +
  labs(y = 'Number of Marriages / Log(Number of Marriages)', x = 'Date')
```
