---
title: "Forecasting Monthly Marriages"
author: "Carlota Marto 20241729, Francisca Teixeira 20241702, Inês Caseiro 20241780"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(forecast)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tsibble)
library(feasts)
library(fable)
library(dplyr)
library(knitr)
library(urca)
```

## 1. Introduction

This project analyzes and forecasts **monthly marriages in Queensland, Australia**, using the Box-Jenkins methodology (ARIMA/SARIMA).\
The dataset is sourced from the Queensland Government Open Data portal ([Marriages by Month](https://www.data.qld.gov.au/dataset/marriages-by-month)), which provides official monthly counts of marriages registered across the state.

The objective of this analysis is to identify trends and seasonal patterns in the number of marriages, fit a suitable SARIMA model, and produce reliable forecasts for the next 12 months.\
These forecasts can provide insights for government planning, service provision, and understanding social trends related to marriage in Queensland.

Session -> Set Working Directory as the directory where this file is currently in the computer.
```{r load-data}
# Define the correct path for the CSV file
file_path <- "marriages_queensland/combined_file.csv"

# Load the CSV file
data <- read.csv(file_path)

# Create a Date column from Year and Month
data <- data %>%
  mutate(Date = yearmonth(paste(year, Month, sep = "-"))) %>%
  arrange(Date)

# Convert to tsibble
data_ts <- data %>%
  select(Date, Marriages) %>%
  as_tsibble(index = Date)

# Show the first 10 rows of the dataset
kable(head(data_ts, 10), caption = "First 10 rows of Monthly Marriages Data")

```

```{r show data plot}
data_ts %>%
  ggplot(aes(x = Date, y = Marriages)) +
  geom_line(color="red") +
  labs(title="Monthly Marriages Over Time", x="Date", y="Marriages") +
  theme_minimal()
```
```{r create train}
data_train <- data_ts%>%
  filter(year(Date) <= 2020)
data_train
```


```{r }
data %>%
  mutate(log_Marriages = log(Marriages)) %>%  # Log transform the 'Marriages' column
  pivot_longer(cols = c(Marriages, log_Marriages), names_to = "Transformation", values_to = "Value") %>%  # Reshape data to long format
  ggplot(aes(x = Date, y = Value)) +  # Plot Date vs the transformed or original 'Marriages'
  geom_line() +
  facet_wrap(~ Transformation, scales = "free_y") +  # Facet by Transformation (original vs log-transformed)
  theme_minimal() +
  labs(y = 'Number of Marriages / Log(Number of Marriages)', x = 'Date')
```

As the variance is not constant over time, we need to stabilize it. For that, we will use the Box-Cox transformation with the guerrero method for finding lambda.
```{r lambda}
lambda <- data_train %>%
  features(Marriages, features = guerrero) %>%
  pull(lambda_guerrero)
lambda
```


```{r applying box-cox transformation}
data_train <- data_train %>%
  mutate(TransformedMarriages = box_cox(Marriages, lambda))

```



# Stationarity
```{r}
data_train %>%
  gg_tsdisplay(y = TransformedMarriages, plot_type = "partial") +
  geom_line(color="red") +
  labs(title="Monthly Marriages Over Time", x="Date", y="Marriages") 

```


```{r apply DF test}
summary(ur.df(data_train$TransformedMarriages,type = "drift" ,lags = 16))

```
As test-statistic (1.5748) does not belong to the critical region (tobs < -2.88), we fail to reject H0. There is statistical evidence of non-stationarity at 5% significance level. /
We will apply a seasonal difference.


```{r seasonal difference}

data_train %>%
  gg_tsdisplay(difference(TransformedMarriages, lag = 12), plot_type = "partial")

```

```{r}
summary(ur.df(na.omit(difference(data_train$TransformedMarriages, lag = 12)), type = "drift", lags = 1))
```
As the test-statistic (-5.0029) belongs to the critical region (tobs < -2.88), we reject H0. There is statistical evidence of stationarity at 5% significance level.

# Model identification

```{r plot 2, warning=FALSE}
data_train %>%
  gg_tsdisplay(difference(TransformedMarriages, lag = 12), plot_type = "partial", lag = 36)

```

Possible models: SARIMA(1,0,3) (2,1,0)12, SARIMA(1,0,1) (0,1,0)12
